---
title: "Overview of the distribution of codes"
output:
  pdf_document:
    toc: yes
  html_notebook:
    fig_caption: yes
    fig_width: 8
    fig_height: 5
    number_sections: yes
    theme: paper
    toc: yes
  word_document:
    toc: yes
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(tidyverse)
require(gridExtra) # for side-by-side plots
require(glue) # for pretty labels
load("../data/codings.rda")
```

These notes describe the distribution of codings attributed by raters to the qualitative data, and raises exploratory explanations. This exploration step excludes examination of such explanations. Other notebooks will examine the possible explanations raised here, for instance through by probing relationships between codes, or extracting exemplary quotations. Similarly, building of a typology will appear elsewhere.

# Data
Analysis is based on the `codings` dataset, in which each row represents a code attributed to a story. It looks like this:

```{r}
glimpse(codings)
```
Where:

- `fid` is the story identifier
- `codename` is a code attributed by a rater, following a \<category\>-\<code\>-\<subcode\> convention.
- `category`, `code`, `subcode` unfold the coding convention
- `char_fid` is the total number of characters in the story
- `char_start` and `char_end` are characters delimiting the excerpt for which the code was attributed
- `description` is the description of codes or subcodes, formulated as an answer to a rating question
- `category.desc` is the description of the category, formulated as a rating question

The pair (`fid`, `codename`) is the primary key. The dataset was created by the `retrieve_data.R` script, which retrieves it from the RQDA database, and merges fields in several tables. No previous transformations were performed.

This analysis will take top-down view, looking first at code categories, and then at codes and subcodes. 

# Categories
## Descriptions

The category descriptions are as follows:
```{r}
codings %>%
  select(Category = category, Description = category.desc) %>%
  distinct() %>% 
  knitr::kable(booktabs = T)# %>% kableExtra::
```

## Factor and list categories
For most code categories, raters were tasked with selecting one code per story, as evidenced by *[Choose one]* in the category description. However, for code categories `process` and `aftermath`, raters were free to choose any number codes, as evidenced by *[Choose all that apply]*. These options translate into two different kinds of categories: *factor categories* and *list categories*.

### Factor categories
Categories for which raters were only able to select one code can map directly to categorical variables, or factors. Each code (or code-subcode if available) represents a level of that factor.

For instance, the category `lock` can be converted into factor, with the following levels:
```{r}
codings %>% 
  filter(category == "lock") %>%
  transmute(levels = codename, description) %>%
  distinct() %>%
  arrange(levels) %>%
  knitr::kable()
```

Some levels correspond to codes (following the category-code convention), and some correspond to subcodes (category-code-subcode). Because the codebook was designed to not allow selection of codes when subcodes were available, there is no overlap.

### List categories
Categories `process` and `aftermath` do not enjoy the same property of being easily transformed into factors. They are, instead, list variables. For each story, each category maps to a list of codes or code-subcodes which may have been chosen by raters. The list can be empty or have as many elements as there are options.

For instance, stories were often rated with more than one code in the `process` category. Story 2, for instance, was attributed 5 `process` codes.

```{r}
codings %>% 
  filter(category == "process", fid == 2) %>%
  select(fid, category, code, subcode) %>%
  knitr::kable()
  
```

Because the codes represent processes that occured in an observational unit, it makes sense to think of them as a list.

## Temporal structure
Since stories were collected through a narrative device, it is expectable that code categories follow a temporal pattern. 

The dataset contains an indication of the first character associated with a coding, as `chart_start`, and the total story length, as `char_fid`. From those, a relative location in the text (from 0 to 1) where codes were assigned can be calculated. Categories can then be ordered by the mean relative location where codings started:

```{r}
categories_ordered = 
  codings %>% 
    mutate(
      relative_start = char_start / char_fid
      ) %>%
    group_by(category) %>%
    summarise(
      mean_relative_start = mean(relative_start)
      ) %>%
    arrange(mean_relative_start)
categories_ordered %>%
  knitr::kable()
```

This ordering can be applied to the category variable, for subsequent visualizations.
```{r}
codings = 
  codings %>%
  mutate(category = factor(category, levels = categories_ordered$category))
```

Instead of looking just at the mean, we can also see the distribution of relative start characters across categories, for instance with a violin plot:

```{r}
codings %>% 
  mutate(relative_start = char_start / char_fid) %>%
  ggplot +
  aes(
    x = category, 
    y = relative_start
    ) + 
  geom_violin(scale = "width", trim = F, draw_quantiles = c(0.5)) +
  geom_point(alpha = 1/5)
```
Violins are not very compact. An alternative, which may be more suitable for publication, is a vertical point range plot.

```{r}
codings %>% 
  mutate(relative_start = char_start / char_fid) %>%
  ggplot +
  aes(
    x = category, 
    y = relative_start
    ) + 
  stat_summary(
    fun.y = mean,
    fun.ymax = max,
    fun.ymin = min,
    geom = "pointrange"
  ) +
  geom_point(alpha = 1/5) + 
  coord_flip()

```

*Observations:*

The temporal structure does reflect a narrative structure, and matches our observations of how the stories usually progressed, such as the following fictitious example:

> Ash and Val were friends (`relationshiptype`). Ash went to the bathroom and left the phone behind (`opportunity`). Val wanted to play a prank (`motivation`). The phone did not have a lock (`lock`), so Val got in and took stupid pictures (`process`). Ash found out a few days later (`knowledge`), was very mad (`aftermath`), and never spoke to Val again (`status`). 

To some extent, this structure might have been super-imposed in the story editing process. Coded raw stories can be checked for comparison.

# Code & subcode distribution
Given the number of codes and subcodes available to raters, visualizing them in a single graph is not ideal. For instance, a bar chart, even without the labels, is very dense:

```{r}
codings %>%
  ggplot() +
  geom_bar(
    mapping = aes(
      x = category,
      fill = codename
    ),
    show.legend = F,
    alpha = 0,
    color = "black"
  ) +
  coord_flip()

```

Some observations, however, still can be drawn. 

1. There is a variety of code distributions within categories. Some codes or subcodes have very few observations, while others have many.

2. As expected, all categories, except for `process` and `aftermath`, have exactly has many observations as there are observational units. Category `process` has a higher number, because it was common that more than one process was involved in an intrusion story. Category `aftermath` has fewer observations, because many stories did not express emotional consequences to the level required by the coding criteria.

We'll next look at the distribution of codes and subcodes per category. Where applicable, codes will be shown both as raters saw them, and consolidated into codes.

```{r}
# barplots will be created with the same functions (below) for consistency.
get_barplot_grid = function(data_subset, main_size = 3/4) {
    cowplot::plot_grid(
      get_barplot_codenames(data_subset),# + cowplot::theme_cowplot(),
      get_barplot_aggregated(data_subset), #+ cowplot::theme_cowplot(),
      nrow = 2,
      rel_heights = c(main_size, 1-main_size),
      align = "v"
    )
}

get_barplot_aggregated = function(data_subset) {
  data_subset %>% 
    ggplot() + 
    geom_bar(
      aes(
        x = fct_infreq(code),
        fill = code
      ),
      show.legend = F
    ) +
    coord_flip() +
    xlab(NULL) + 
    ylab("Consolidated frequency")
}

get_barplot_codenames = function(data_subset, width = 80) { 
  data_subset %>% 
  mutate(
    variable = glue(
      "[{codename}] \n {wraped_desc}",
      wraped_desc = str_wrap(description, width = width)
      )
    ) %>%
  ggplot() + 
    geom_bar(
      aes(
        x = fct_infreq(variable),
        fill = code
      ),
      show.legend = F
    ) +
    coord_flip() +
    xlab(NULL) + 
    ylab("Rater decision frequency") +
  theme(
    axis.text.y = 
          element_text(
            size = 10, 
            hjust = 0
            )
    )
}
```

## Category `relationshiptype`
The following types of relationship were coded:

```{r, fig.width=8}
codings %>%
  filter(category == "relationshiptype") %>%
  get_barplot_grid() 

```
*Observations:*

- In most stories, the relationship between subjects was coded as `intimate`. 

- Of the non-intimate type of relationship subcodes, `friends` was much more commonly selected than other available types, namely `acquaintance`, `family`, or `professional`.

- It is conceivable that relationships being intimate or non-intimate is a preponderant factor on how stories progress and their eventual outcomes. Relationships between these codes, and all subsequent codes, could reveal the relationship type as very significant dimension.

- Non-intimate types of relationships were distinguished by subcodes, while `intimate` was a code without subcodes. Subcodes indicating whether it is a current, former, or aspirational type of relationship, has the code description suggests, could be added, if there's any indication that such distinctions could have an impact on other aspects of stories.

## Category `opportunity`
The following types of intrusion opportunities were coded:

```{r fig.width=8, fig.height=7}
codings %>%
  filter(category == "opportunity") %>%
  get_barplot_grid()
```

*Observations:*

* Raters judged that intrusions happened, in most stories, when devices were left unattended by the passive subject. 

* Cases where intrusions happened after devices were voluntarily handed to the active subject (which fall under the `deception` code) were rare. This difference, if reflective of a wider reality, suggests that security measures for devices at rest may be more impactful than security measures for addressing impromptu sharing scenarios. Several forces may be conspiring to make sharing scenarios less favorable for intrusions, including social etiquette, the probability of getting caught, and the hassle to successfully deceive the passive subject. It is worth revisiting the relationship between use of deception, and the motivation, processes, and aftermath of the intrusion -- it may be the case that deception is associated with the least high-stakes cases, where getting caught is of no concern and does not breach social norms; or it may be the case that it is associated with the most high-stakes cases, where the hassle and risk of getting caught are worth the trouble, from the active subject's perspective.

* Raters found several scenarios where devices are left temporarily unattended, expressed as subcodes. Most commonly, they found stories to indicate devices being unattended while their owners went to the bathroom (subcode `unattended-bathroom`). Raters also found stories where devices were left unattended in locations often deemed "trusted", such as the home or office, while the passive subject went somewhere else, for instance to run errands, or to a meeting (subcodes `unattended-outside`, and `unattended-meeting`); and stories where passive subjects were asleep (`unattended-asleep`). These stories of intrusions thus substantiate the questioning of common assumptions of trustworthiness associated with locations.

* The commonality of the "bathroom scenario" seem to suggest that intrusions can occur in very little time. This aspect might be of importance for system design. For instance, intrusion detection algorithms which require a substantial period of behavior acquisition prior to classification may be ineffective in practice.

* It can also be speculated that the "bathroom scenario" provides a set of relatively safe, or otherwise favorable, conditions for active subjects. Such conditions may include a *predictable* length of time alone with the device, *forewarning* that the passive subject may return (e.g. aural notice of flushing, shower shutting off, or doors opening), *social etiquette* in favor of leaving the device unattended (e.g. so as not to signal distrust), and passive subject's *trust* they are in a safe environment.

* Situations where the opportunity for access came through access to a secondary device (code `opportunity-secondarydevice`) were not common, but still existed. These cases are also worth examining in comparison to unattended access, as in both cases there is a low probability of getting caught.


## Category `motivation`
The following types of motivation for intrusion were coded:

```{r, fig.width=8, fig.height=8.5}
codings %>%
  filter(category == "motivation") %>%
  get_barplot_grid()
```

*Observations:*

* "Control" was defined as the passive subject wanting "to learn about, or influence" relationships between the passive subject and third parties. By a clear majority, codes related to control were the most selected by raters, suggesting that these is the most common motivation for intrusion among parties known to each other.

* A desire to control possible intimate relationships of the passive subject (code `control-intimate`) is the most commonly selected subcode. Research on technology-mediated intimate partner abuse has previously identified control as an important, or even principal motivation (although often "control" is construed more broadly). Although definitions of abuse vary, our observation when reading the stories was that there were few examples that would undoubtedly qualify as intimate partner abuse. 

* Nevertheless, this evidence does seem to lend support a more modest claim, that intrusions to smartphones can be a component of controlling behaviors between intimate partners. Furthermore, this data suggestd that technology-enabled controlling behaviors between intimate partners may be better understood as existing on a continuum, and not necessarily under a dichotomy between abusive vs. healthy. 

* Other motivations which raters coded were using some functionality of the device (code `convenience`), playing pranks (code `prank`), and exploitation (`exploit`).

* Exploitation was rated with subcodes that refer to behaviors that are commonly seen as extremely pernicious, namely stealing of devices (code `exploit-steal`), stealing of business information (`exploit-business`), or stealing of sexualizable media content (`exploit-sexploit`).

* These exploitative motivations often deserve a great deal of attention, both in research and in media. Alarm over these types of occurrences has informed interventions such as security technology, security advice, public policy, and business investment. In contrast, control-motivated intrusions, which do appear in the data much more commonly, have at best received less attention, and at worst been regarded as unimportant or unaddressable. We can only speculate, but a reason behind this asymmetry may be that, in control-motivated intrusions, it is often more difficult to assign polarizing roles to parties, such as perpetrator vs. victim. For instance, when an active subject seeks confirmation of partner infidelity, and finds it, it may be more difficult to label them "perpetrators of intrusion", and thus to strike enough fear to prompt interventions against those types of intrusions. Interventions, such as security technologies directed at preventing control-motivated intrusions by intimate partners would, conceivably, be seen as enabling behaviors commonly deemed censurable, such as intimate partner infidelity.

## Category `lock`
The following types of locking existence and effectiveness were coded:

```{r, fig.width=8}
codings %>%
  filter(category == "lock") %>%
  get_barplot_grid()
```

*Observations:*

* In most stories, raters judged that mentions to either presence or effectiveness of locks were not present. The codebook instructed raters to not make inferences, but to code only explicit evidence.

* Raters judged there was evidence of locks being effective in small subset of stories. An effective lock, however, does not necessarily prevent all kinds of intrusions. For instance, pre-lock notifications may be accessible. These stories must thus be inspected to check whether they should be excluded from analysis.

* In few stories, raters found evidence that locks were not set (code `notset`). 

* More commonly, however, raters found evidence indicating locks were set, but active subjects were still able to obtain unauthorized access (code `ineffective`). This lends support to questioning the effectiveness of locking as an effective defense against socially-close adversaries.

* The subcodes further qualify these observations for different types of authentication methods. Subcodes `ineffective-known`, `ineffective-observed`, and `ineffective-easy`, indicate that locks were ineffective due to known constraints of secret-based authentication methods, respectively password-sharing practices, shoulder-surfing, and informed guessing. In contrast, only subcode `ineffective-unlocked` may apply to biometric authentication (inspection of codings should clarify it that was the case). One possible implication of this observation, is that adoption of biometric unlock methods may greatly mitigate the possibility of intrusions by socially-close adversaries; and recommending its adoption a sensible advice for many users.

## Category `process` (list)

> TRANSFORMATION: "snooping" was renamed to "information gathering". The code name "snooping" is more informal than the remaining code names in the process category, i.e. "tampering", "exfiltration", and "impersonation". Since it's only a label for the description raters actually used, there is no problem in changing after the fact.

```{r echo=T}
codings = 
  codings %>% 
  mutate_at(vars(codename, code), gsub,  pattern = "snooping",
    replacement = "informationgathering")
```


The following types of processes were coded:

```{r, fig.width=8, fig.height=10}
codings %>%
  filter(category == "process") %>%
  get_barplot_grid(main_size = 3.5/4)
```

*Observations:*

* Since several codes could, and were often selected for the same stories, these frequency charts have a different reading. They do not inform on the proportion of stories in which each code or code-subcode was found. It might make sense to produce such charts in subsequent analysis.

* Given the large number of processes which were coded, it could also be useful to  find clusters of co-occurrence. 

*	Raters most often found evidence in the stories of information gathering. Raters found evidence of active subjects inspecting a wide variety of information sources, including media files such as photos, social media activity, contacts, notifications, apps installed, call logs, or internet history. Most commonly, by far wide margin, raters found evidence of active subjects inspecting text-based records of conversations, such as text messages, instant messages, or emails.

*	Solely by frequency, it can be speculated that there is a connection between active subjects being motivated by controlling relationships of the passive subject, and conducting intrusions for information gathering. This claim can be further examined by looking at relationships between these codes. If true, it starts to paint a picture of the most common type of intrusions found in the stories: intimate partners, finding opportunities optimal for avoiding detection, and undeterred by secret-based locks, exerting their desire for control by snooping on their partner’s communications with third parties.


* Raters judged stories to provide evidence of other processes of intrusions, namely tampering with the device (code `tampering`), extracting data to outside media (code `exfiltration`), and impersonating the passive subject in interactions with others (code `impersonation`). 

* It can be speculated that subcodes of `tampering` are associated with with different sets of `motivation` codes. Changing settings (`tampering-settings`) or contacts (`tampering-contacts`), taking photos (`tampering-media`), and even deleting content (`tampering-deletion`) might be associated with several types of intrusions that might not be considered "high-stakes", such as playing pranks or attempting to expediently accomplish some task which the passive subject might not object to. However, implanting surveillance software (`tampering-implant`) appears particularly pernicious, and is a behavioral marker explicitly mentioned in prior research on technology-mediated intimate partner abuse. 

* Subcodes of exfiltration to another medium (of contacts, conversations, and visual media), and the single subcode of impersonation (in conversations with third parties) are also ambiguous on their own, warranting additional examination of co- occurrence with motivations and aftermath, and examination of co- occurrence with other processes.

## Category `knowledge`
The following types of ways in which active subjects learned about intrusions were coded:

```{r, fig.width=8}
codings %>%
  filter(category == "knowledge") %>%
  get_barplot_grid(main_size = 3/4)
```

*Observations:*

* Raters judged that in most stories passive subjects eventually learned about intrusions. With subcodes, raters indicated three paths to knowledge: findings active subjects in the act (subcode `yes-redhanded`), self-admission by active subjects (subcode `yes-admission`), and finding clues leading to a suspicious on unauthorized access (subcode `yes-clues`). Several types of clues are suggested in the subcode description, and some particular examples would illustrate similarities and differences of possible interest.

* The distribution of this code between `yes` and `no` may be an artifact of the data-gathering. It is reasonable to assume that active subjects are likely to under-report, and thus most stories could have been provided by passive subjects, which can only tell them if they eventually learned about intrusions. However, as our Facebook list experiment indicated, there is a comparable bias on both the active and the passive subject sides. Thus, it might be true that eventually, a larger portion of intrusions that people remember are the ones where eventually the passive subject learned of the intrusions.

## Category `aftermath` (list)

The following types of explicitly-referenced sentiments towards the intrusion were coded:

```{r, fig.width=8}
codings %>%
  filter(category == "aftermath") %>%
  get_barplot_codenames()
```

*Observations:*

* Sentiments were rated mostly as a control variable for perceptions of severity -- at the outset, we planned a second study were stories would be rated as to their severity, and it is reasonable to expect that readers would let their sentiments be influenced by the characters own sentiments.

* In this category, only codes, and not subcodes, were attributed. There is a natural structure between passive and negative types of aftermath, as evidenced in the code descriptions. The data could be transformed to superimpose this distinction as a code, but for now the bar charts seems illustrative enough.

* Nevertheless, we can observe that, in most cases, raters judged intrusions to lead to negative sentiments, such as annoyment, anger, guilt, humiliation, pain, etc., on both the active and the passive subjects. In some cases, positive sentiments arose, such as amusement, satisfaction, or relief. 

* Examining relationships of aftermath codes with prior categories should inform the origins and context of this distribution.

* Like in the case of process codes, these charts are not relative to the proportion of stories, and subsequent analysis, for instance clustering, should address that limitation, and re-open examination of these observations.

## Category `status`
The following types of resulting relationship status were coded:

```{r}
codings %>%
  filter(category == "status") %>%
  get_barplot_codenames() 
```

*Observations:*

*	Resulting relationship status was coded, like aftermath, mostly as a control variable for perceptions of severity.

*	Most commonly, stories were not coded as evidencing maintenance or change in relationship status, per the coding criteria. When status was rated, both maintenance and ending of relationships was found.

*	Like aftermath, this category can be seen as an outcome variable, and could analyzed as being predicted by preceding code categories.

```{r echo=T}
# TODO:
# - Move mid-notebook transformations in `process`subcodes somewhere else, e.g. RQDA tables or retrieval script.
#- RQDA: Re-inspect codings in RQDA.
#- RQDA: Correct typos in category and code descriptions
#- Consider coding whether the passive subject was engaged in marital or dating infidelity.  This dimension could affect perceptions of severity. This could be done after regression analysis of severity ratings, if there indication that adding this variable could reduce variance.
```


# TODOs

