---
title: "Model simulation"
output: html_notebook
---

This notebook contains simulations of approaches to modelling, to inform the design of study 2.

First, load the data.

```{r include=FALSE}
library(here)
library(tidyverse)
library(modelr)
library(ordinal)
load(here("data", "codings.rda"))
set.seed(1980)
```

Suppose that there was just one measure of severity, and that and a single rater classified all stories as to their severity. The simplest model would take the form:

> severity ~ story  + error

The measure of severity could be a continous-approaching rating, ordinal, or binary. Random severity outcomes can be simulated:

```{r echo = T, include = FALSE}
n = codings %>%
  distinct(fid) %>%
  nrow()

# continuous severity from 0 - 100
severity_num = 
  runif(n) * 100

# binary
severity_binary =
  sample(0:1, n, replace = T) == 0

# 5 levels of an ordered factors
levels = c("not_severe", 
           "slightly_severe",
           "somewhat_severe",
           "moderately_severe",
           "extremely_severe")
severity_factor =
  sample(levels, n, replace = 2) %>%
  factor(levels = levels, ordered = T)

```

Since these are randomly generated outcomes, completely unrelated to the predictor (story), the models should not be informative.

```{r}
# fid as categorical
story = codings$fid %>% unique() %>% factor(ordered = F)

# fit models
linear = lm(severity_num ~ story)
logistic =  glm(severity_binary ~ story,
                family = "binomial")
ordinal = clm(severity_factor ~ 1)

# model predictions 
tibble(story, severity_num, severity_binary) %>%
  add_predictions(
    model = linear, 
    var = "severity_num_pred") %>%
  add_predictions(
    model = logistic,
    var = "severity_binary_pred") %>%
  select(order(names(.))) %>%
  select(story, everything())
```

Alas, the linear model becomes a map from story -> rating, thus perfectly predicting the response. This happens because `story` is a categorical variable. Predictions from the logistic model takes only two values, positive or negative.

```{r}
codings %>% 
  group_by(story = fid) %>% 
  summarise(storylength = first(char_fid)) %>%
  mutate(
    severity_num,
    severity_binary
  ) %>%
  add_predictions(
    lm(severity_num ~ storylength, data = .)
  )

story = codings$fid %>% unique() %>% factor(ordered = F)

# fit models
linear = lm(severity_num ~ story)
logistic =  glm(severity_binary ~ story,
                family = "binomial")

# model predictions 
tibble(story, severity_num, severity_binary) %>%
  add_predictions(
    model = linear, 
    var = "severity_num_pred") %>%
  add_predictions(
    model = logistic,
    var = "severity_binary_pred") %>%
  mutate(
    severity_binary_pred = severity_binary_pred > 0
  ) %>%
  select(order(names(.))) %>%
  select(story, everything())


```

