---
title: "Sample summaries for 1st study participants"
output: html_notebook
---
```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "..")
source("../R/retrieve_data.R")
```
## Data retrieval
Participant data, exported from LimeSurvey, is stored in Google Sheet. 

Start by retrieving the data. The key to the appropriate Google Sheet is encrypted in the vault. Out-of-band authentication may be needed to
access Google Sheets.

```{r, results='hide'}
data = retrieve_sheet_data("s1_participants")
```
Here are the available variables per participant:
```{r}
names(data)

```

## Sample summaries
### Gender

```{r}
data %>% group_by(GENDER) %>% summarise(n()) 
```
### Age group
```{r}
data %>% group_by(AGE) %>% summarise(n()) 
```
### Country
```{r}
data %>% group_by(country) %>% summarise(n()) 
```
### Wave of data collection
```{r}
data %>% group_by(wave) %>% summarise(n())
```
### Data gathering period
First and last answer submission data/time.
```{r}
data %>% summarise(first = min(submitdate), last = max(submitdate)) 
```
### Task completion time
Mean time to complete the survey-answering task, in minutes, and respective SD.
```{r}
data %>% 
  mutate(tasktime.m = interviewtime / 60) %>%
  summarise(mean(tasktime.m), sd(tasktime.m))
```
### Compensation
Mean hourly compensation, in GBP, and respective SD.
```{r}
data %>%
  mutate(pph = 60 * 60 * compensation / interviewtime) %>%
  summarise(mean(pph), sd(pph))
```
### Missed ACQs
Instances where at least one attention check question was missed.
```{r}
data %>% 
  filter(QUAL1 != "Yes" | QUAL2 != "Yes") %>%
  mutate(tasktime.m = interviewtime / 60) %>%
  select(id, wave, tasktime.m)
```
Interview times look normal. Relayed stories were manually checked, and although not the best, judged credible and informative. 

