---
title: "Sample summaries for study participants"
output: html_notebook
---
```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

> This notebook is not reproducible without access to raw data.

## Data retrieval
Participant data, exported from LimeSurvey, was stored in Google Sheet, and retrived to a local file in `data-raw`. Start by retrieving the data. 

```{r, results='hide'}
data = read_csv(here("data-raw/retrieved_participants.csv"))
```

Here are the available variables per participant:

```{r}
names(data)

```

## Sample summaries
### Gender

```{r}
data %>% group_by(GENDER) %>% summarise(f = n()) 
```

### Age group
```{r}
data %>% group_by(AGE) %>% summarise(f = n()) 
```

### Country

Country was inferred from IP address, using the Maxmind database.

```{r}
data %>% group_by(country) %>% summarise(f = n()) %>% arrange(desc(f))
```
### Waves of data collection
```{r}
data %>% group_by(wave) %>% 
  summarise(f = n(), 
            day = min(submitdate))
```

### Task completion time
Mean time to complete the survey-answering task, in minutes, and respective SD.
```{r}
data %>% 
  mutate(minutes = interviewtime / 60) %>%
  summarise(mean = mean(minutes), 
            sd = sd(minutes))
```
### Compensation
Mean hourly compensation, in GBP, and respective SD.
```{r}
data %>%
  mutate(pph = 60 * 60 * compensation / interviewtime) %>%
  summarise(mean = mean(pph), 
            sd = sd(pph))
```
### Missed ACQs
Instances where at least one attention check question was missed, but data was not excluded. Relayed stories were manually checked, and judged credible and informative. 

```{r}
data %>% 
  filter(QUAL1 != "Yes" | QUAL2 != "Yes") %>%
  select(id, wave, QUAL1, QUAL2)
```


