---
title: "Story lenght statistics, pre- and post-editing"
output: html_notebook
---
```{r echo = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

> This notebook is not reproducible without access to raw data.

For stories 1 to 35, there are both raw and edited versions. We can compare word counts, and see if there are significant differences. Since many stories were edited for brevity, some difference is expected.

Start by retrieving the data:

```{r echo = T, results = "hide"}
stories = 
  read_csv(here("data-raw/stories.csv"))
```

Now, add word counts to stories which have both raw and edited versions. 

```{r}
word_count =
  stories %>%
  filter(! is.na(edited_story)) %>%
  mutate(raw_wc = str_count(raw_story),
         edited_wc = str_count(edited_story)) %>%
  select(fid, raw_wc, edited_wc)
glimpse(word_count)
```

Calculate means, SDs:

```{r}
word_count %>%
  summarise(
    mean_wc_edited = mean(edited_wc),
    sd_wc_edited = sd(edited_wc),
    mean_wc_raw = mean(raw_wc),
    sd_wc_raw = sd(raw_wc)
    )
```

Edited stories, on average, indeed have less words.

With boxplot:

```{r}
word_count %>%
  gather(raw_wc, edited_wc, key = "type", value = "wc") %>%
  ggplot(data = .) +
  geom_boxplot(aes(type, wc))
```

Medians are very close; most of the reduction in word count comes for outliers.

Are word counts significantly different?

```{r}
t.test(word_count$raw_wc, word_count$edited_wc, paired = T)
```

