---
title: "Visualizing codings with diagrams"
output: html_notebook
---
```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "..")
source("setup.R")
codings = getCodes()
```

These are notes on visualizing codes with diagrams. This is a step in constructing a typology of instances of unauthorized access.

# Objectives
The objectives of this analysis are as follows:

1. Visualize code frequencies in each category.
2. Visualize relationships between codes of different categories.

# Issues and observations prior to analysis
## Temporal pattern
As is expected by data collections through a narrative device, the code categories follow a temporal pattern. Centering stories around the instance or pattern of unauthorized access, the codes reflect three periods in time: *before*, *during*, and *after*.

1. Before: code categories *relationshiptype* and *motivation*.
2. During: code categories *opportunity*, *lock*, and *process*.
3. After: code categories *knowledge*, *aftermath*, and *status*.

See the [codefrequency](codefrequency.nb.html) notebook for descriptions of categories. 

> TODO: represent temporal pattern

## Level of analysis
For this analysis, the focus will be on codes, *not* subcodes. 

The codings data looks like this:
```{r}
str(codings, vec.len = 1, ncharmax = 80)
```

To analyze at the code level, where only category and code are needed, only one instance of code per story is needed.

```{r}
codings.wo.subcodes =
  codings %>% 
  select(fid, category, code) %>%
  distinct()
str(codings.wo.subcodes, vec.len = 1, ncharmax = 80)
```

## Dealing with multi-code categories
For most code categories, only one code is selected per observational unit (i.e., fid). However, for code categories *process* and *aftermath*, raters were free to not choose any code, or to choose many. Thus, each code in on of these categories can be seen as a category itself, with a binary value. This fact presents challenges to subsequent analysis, which will often require exactly 1 value per observational unit.

To be able to proceed with exploring visualizations, an additional dataset is created limited to the remaining categories. To easily conform with frequency analysis, categories are spread to collumns.

```{r echo=TRUE}
data.wide = 
  codings.wo.subcodes %>% 
  filter(category != "process" & 
           category != "aftermath") %>%
  spread(category, code) %>%
  # order by timeline
  select(fid, relationshiptype, motivation, opportunity, lock, knowledge, status)
str(data.wide, vec.len = 2)
```

Another common data representation used in visualizations is as a frequency table, where each line contains the frequency of a given combinations of codes.

```{r}
data.wide.freq =
  data.wide %>%
  select(-fid) %>%
  group_by_all() %>%
  summarise(freq = n()) %>%
  ungroup() %>%
  arrange(desc(freq))
str(data.wide.freq, vec.len = 3)
```

### Transformation

> TODO: revise when additional data comes in

Another approach that could be considered would be to transform the data in multi-code categories, so that they comply with the 1 observation per story format.

For both multi-code categories, cases where raters did not select any code can be trivially transformed in a "unspecified"-type level.

The problematic cases are those for which there was more than one coding. 

For aftermath, these are the cases with multiple codings:

```{r}
codings.wo.subcodes %>%
  select(fid, category, code) %>%
  filter(category == "aftermath") %>%
  arrange(code) %>%
  group_by(fid) %>%
  nest(code) %>%
  mutate(n = purrr::map_int(data, nrow),
         codes = purrr::map(data, unlist),
         codelist = purrr::map_chr(codes, paste, collapse = ", ")) %>%
  filter(n > 1) %>%
  select(fid, codelist) %>%
  arrange(codelist)
```
There appear to be the following patterns:

- Active subject has mixed feelings
- Both subjects have negative feelings

There is one instance which does not conform to this patter, which could be labeled as "others".

For process, these are the cases with multiple codings:

```{r}
codings.wo.subcodes %>%
  select(fid, category, code) %>%
  filter(category == "process") %>%
  arrange(code) %>%
  group_by(fid) %>%
  nest(code) %>%
  mutate(n = purrr::map_int(data, nrow),
         codes = purrr::map(data, unlist),
         codelist = purrr::map_chr(codes, paste, collapse = ", ")) %>%
  filter(n > 1) %>%
  select(fid, codelist) %>%
  arrange(codelist)
```
It appears that all combinations include snooping, plus one other kind of code.

However, it is not true that those other codes subsume "snooping". They sometimes occur on their own:

```{r}
snooping.fids =
  codings.wo.subcodes %>% 
  filter(category == "process" & code == "snooping") %>%
  .$fid

codings.wo.subcodes %>%
  filter(category == "process" &
           ! fid %in% snooping.fids)

```

A rule for transformation could be always choosing the other code. In that case, any selection of on of the other codes should be described as *possibly* subsumming "snooping". One disadvantage to this approach would be to possibly distort an analysis of the relationshiip of "snooping" codes and other codes.


# Alluvial diagrams 
## With [alluvial](https://CRAN.R-project.org/package=alluvial)
First, an attempt at creating an alluvial using the "alluvial" package. This package uses standard plotting. 

```{r echo=TRUE}
library(alluvial)
```



Here is the default alluvial representation:

```{r echo=TRUE}
alluvial(
  alluvial.data[,1:6],
  freq = alluvial.data$freq
)
```

The graph conveys the correct code frequencies for each category, limited to the single-choice categories. However, the relationships between codes are difficult to make sense of.

Using this package, the input is a contigency table, and the alluvia can be followed from beggining to end, their thickness representing frequency. The graph would be easier to understand if alluvials only represented pairwise relationships between adjoining categories, i.e. pairwise "of which" relationships. 

Even with these constraints, there is some room to improve on the graph, by adding colors and tinkering with ordering. 

```{r}
# Create orderings
alluvial.data = 
  alluvial.data %>% 
  mutate(
    relationshiptype.order = 
      case_when(
        relationshiptype == "intimate" ~ 1,
        relationshiptype == "nonintimate" ~ 2,
        relationshiptype == "unspecified" ~ 3,
        TRUE ~ 0
        ),
    motivation.order = 
      case_when(
        motivation == "control" ~ 1,
        motivation == "exploit" ~ 2,
        motivation == "prank" ~ 3,
        motivation == "convenience" ~ 4,
        motivation == "accidental" ~ 5,
        motivation == "unspecified" ~ 6,
        TRUE ~ 0
        ),
    opportunity.order = 
      case_when(
        opportunity == "unattended" ~ 1,
        opportunity == "deception" ~ 2,
        opportunity == "secondarydevice" ~ 3,
        opportunity == "unspecified" ~ 4,
        TRUE ~0
      )
  ) 
alluvial(
  alluvial.data[,1:6],
  freq = alluvial.data$freq,
  border = alluvial.data$relationshiptype.order + 1,
  col = alluvial.data$relationshiptype.order + 1,
  ordering = list(
    alluvial.data$relationshiptype.order,
    NULL,
    NULL,
    NULL,
    NULL,
    NULL
  ),
  gap.width = .1,
  cex = .7,
  blocks = T
  #alpha = 0.7,
)
```


Code for the previous graph is very hacky. The vertical ordering of levels is an experimental feature, and I found no way of properly ordering past the first category. 

Conclusion: objective 1, of representing code frequencies, is partially accomplished, by the stacked barplot format of the alluvial nodes. However, relationships of adjoining codes are difficult to discern. Additionally, there is very little flexibility for tinkering with the basic plot.

## With [ggalluvial](https://CRAN.R-project.org/package=ggalluvial)

> Ongoing

First attempt, using the same data.
```{r}
library(ggalluvial)
alluvial.data %>% 
  ggplot(aes(weight = freq, 
             axis1 = relationshiptype, 
             axis2 = motivation,
             axis3 = opportunity,
             axis4 = lock,
             axis5 = knowledge,
             axis6 = status)) +
  geom_alluvium(aes(fill = motivation), width = 1/12) +
  geom_stratum(width = 1/12, fill = "black", color = "grey") +
  geom_label(stat = "stratum", label.strata = TRUE) +
  scale_x_continuous(breaks = 1:6, labels = c(names(data))[1:6]) + #, labels = names(data)) +
  ggtitle("Story codings alluvial")

```


# Code relationships within moments


