---
title: "Analysis of codebook growth"
output: html_notebook
---
```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "..")
source("setup.R")
library(RQDA)
```
## Preparation
Retrieve to *codings* dataset from RQDA:
```{r}
codings = getCodes()
str(codings, vec.len = 1, nchar.max = 80)
```
Add collumn indicating wave of data gathering:

```{r}
codings = 
  codings %>% 
  mutate(wave = case_when(
                fid <= 35 & fid > 25  ~ "Wave 3",
                fid <= 25 & fid > 13  ~ "Wave 2",
                fid <= 13 & fid > 3  ~ "Wave 1", 
                fid <= 3 ~ "Pilot"
                )
  )
str(codings, vec.len = 1, nchar.max = 80)  
```

> TODO: this will break with new data. Should retrieve wave information from data source, and not hard code it (see participants notebook).

## New codes / subcodes per wave
### Codes
```{r}
codings %>% 
  select(fid, wave, category, code) %>% 
  group_by(category, code) %>% 
  summarise_all(min) %>% 
  arrange(fid) %>% 
  ungroup() 
```
### Subcodes
```{r}
codings %>% 
  select(fid, wave, codename) %>% 
  group_by(codename) %>% 
  summarise_all(min) %>% 
  arrange(fid) %>% 
  ungroup() 
```
### Plots
(Plotting code is messy, and not shown here. See Rmd for details.)

> TODO: plots will also break with new data

```{r, include=F}
# helpers to calculate, for each file, codes/subcodes used so far
cum.subcodes = function(curfid, data) {
  data %>% filter(fid <= curfid) %>% summarise(n_distinct(category.code.subcode)) %>% first()
}
cum.codes = function(curfid, data) {
  data %>% filter(fid <= curfid) %>% summarise(n_distinct(category.code)) %>% first()
}

# get tables of cummulative codes/subcodes
subcodes = codings %>% distinct(fid) %>% rowwise(.) %>% 
  mutate(numsubcodes = cum.subcodes(fid, codings)) %>% arrange(fid)
codes = codings %>% distinct(fid) %>% rowwise(.) %>% 
  mutate(numcodes = cum.codes(fid, codings)) %>% arrange(fid)

# plot codebook growth
plotcodes = function(data, ylab ="", main = "", labeloffset = 0.5) {
  data %>% plot(.,  
                    xlab = "Story", 
                    ylab= ylab, 
                    main = main,
                    pch = 1
                )
  text(data[[1]], data[[2]]+labeloffset, labels = data[[2]], cex = 0.6, offset = 10)
  
  # TODO: tidy this up
  text(x = mean(c(0,3.5)), y = 15, labels = "Pilot")
  abline(v = 3.5)
  text(x = mean(c(3.5,13.5)), y = 15, labels = "Wave1")
  abline(v = 13.5)
  text(x = mean(c(13.5,25.5)), y = 15, labels = "Wave2")
  abline(v = 25.5)
  text(x = mean(c(25.5, 35)), y = 15, labels = "Wave3")
}

```
```{r}
plotcodes(subcodes, ylab= "# codes / subcodes used", "Cummulative codes / subcodes", labeloffset = 1)
```
```{r}
plotcodes(codes, ylab= "# codes used", "Cummulative codes")
```


