---
title: "Analysis of codebook growth"
output: html_notebook
---
```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "..")
source("setup.R")
library(RQDA)
```
## Preparation
The data for this analysis are codings stored on the RQDA database. We need to open the project, retrieve the data, than tidy it up.

Retrieve to *codings* dataset:
```{r}
openProject(RQDA_PROJECT_PATH)
codings = getCodingTable()
closeProject()
str(codings)
```
Tidy-up:

> TODO: this will break with new data

```{r}
# create matrix of cat / code / subcode
codings.matrix = 
  str_split_fixed(codings$codename, "-", 3) %>% 
  as_data_frame()
colnames(codings.matrix) = 
  c("category", "code", "subcode")
# bind them
codings = 
  codings %>% 
  bind_cols(codings.matrix)
# add wave 
codings = 
  codings %>% 
  mutate(wave = case_when(
                fid <= 35 & fid > 25  ~ "Wave 3",
                fid <= 25 & fid > 13  ~ "Wave 2",
                fid <= 13 & fid > 3  ~ "Wave 1", 
                fid <= 3 ~ "Pilot"
                )
  )
# tidy-up
codings = 
  codings %>% 
  mutate(category.code = paste(category, code, sep="-"),
         category.code.subcode = codename) %>%  
  select(fid, wave, category.code, category.code.subcode)
str(codings)
```
(This is a bit ugly, but works. See alluvials notebook for more elegant approach.)

## New codes / subcodes per wave
### Codes
```{r}
codings %>% 
  select(fid, wave, category.code) %>% 
  group_by(category.code) %>% 
  summarise_all(min) %>% 
  arrange(fid) %>% 
  ungroup() 
```
### Subcodes
```{r}
codings %>% 
  select(fid, wave, category.code.subcode) %>% 
  group_by(category.code.subcode) %>% 
  summarise_all(min) %>% 
  arrange(fid) %>% 
  ungroup() 
```
### Plots
(Plotting code is messy, and not shown here. See Rmd for details.)

> TODO: this will break with new data

```{r, include=F}
# helpers to calculate, for each file, codes/subcodes used so far
cum.subcodes = function(curfid, data) {
  data %>% filter(fid <= curfid) %>% summarise(n_distinct(category.code.subcode)) %>% first()
}
cum.codes = function(curfid, data) {
  data %>% filter(fid <= curfid) %>% summarise(n_distinct(category.code)) %>% first()
}

# get tables of cummulative codes/subcodes
subcodes = codings %>% distinct(fid) %>% rowwise(.) %>% 
  mutate(numsubcodes = cum.subcodes(fid, codings)) %>% arrange(fid)
codes = codings %>% distinct(fid) %>% rowwise(.) %>% 
  mutate(numcodes = cum.codes(fid, codings)) %>% arrange(fid)

# plot codebook growth
plotcodes = function(data, ylab ="", main = "", labeloffset = 0.5) {
  data %>% plot(.,  
                    xlab = "Story", 
                    ylab= ylab, 
                    main = main,
                    pch = 1
                )
  text(data[[1]], data[[2]]+labeloffset, labels = data[[2]], cex = 0.6, offset = 10)
  
  # TODO: tidy this up
  text(x = mean(c(0,3.5)), y = 15, labels = "Pilot")
  abline(v = 3.5)
  text(x = mean(c(3.5,13.5)), y = 15, labels = "Wave1")
  abline(v = 13.5)
  text(x = mean(c(13.5,25.5)), y = 15, labels = "Wave2")
  abline(v = 25.5)
  text(x = mean(c(25.5, 35)), y = 15, labels = "Wave3")
}

```
```{r}
plotcodes(subcodes, ylab= "# codes / subcodes used", "Cummulative codes / subcodes", labeloffset = 1)
```
```{r}
plotcodes(codes, ylab= "# codes used", "Cummulative codes")
```


