---
title: "Sample summaries for 1st study participants"
output: html_notebook
---
```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "..")
source("setup.R")
```
## Data retrieval
Participant data, exported from LimeSurvey, is stored in Google Sheet. 

First, check if Google Sheets are accessible, by running gs_ls() once.
```{r}
gs_ls() %>% is_tibble()
```
The key to the sheet is stored in the encrypted vault. Here it is how it was stored:
```{r, eval=FALSE, results='hide'}
#key = gs_ls() %>% 
#  filter(grepl("1RUDyCr6C", sheet_key)) %>% 
#  select(sheet_key) %>%
#  first()
# encrypt(.name = SHEETS_VAULT, s1_participants = key)

```
Retrieve the key and read the data:
```{r, results='hide'}
key = decrypt(SHEETS_VAULT)$s1_participants
data = gs_read(gs_key(key))
```
Here's the collumns in the dataset:
```{r}
names(data)
```
## Sample summaries
### Gender
```{r}
data %>% group_by(GENDER) %>% summarise(n()) 
```
### Age group
```{r}
data %>% group_by(AGE) %>% summarise(n()) 
```
### Country
```{r}
data %>% group_by(country) %>% summarise(n()) 
```
### Wave of data collection
```{r}
data %>% group_by(wave) %>% summarise(n())
```
### Data gathering period
First and last answer submission data/time.
```{r}
data %>% summarise(first = min(submitdate), last = max(submitdate)) 
```
### Task completion time
Mean time to complete the survey-answering task, in minutes, and respective SD.
```{r}
data %>% 
  mutate(tasktime.m = interviewtime / 60) %>%
  summarise(mean(tasktime.m), sd(tasktime.m))
```
### Compensation
Mean hourly compensation, in GBP, and respective SD.
```{r}
data %>%
  mutate(pph = 60 * 60 * compensation / interviewtime) %>%
  summarise(mean(pph), sd(pph))
```
### Missed ACQs
Instances where at least one attention check question was missed.
```{r}
data %>% 
  filter(QUAL1 != "Yes" | QUAL2 != "Yes") %>%
  mutate(tasktime.m = interviewtime / 60) %>%
  select(id, wave, tasktime.m)
```
Interview times look normal. Relayed stories were manually checked, and although not the best, judged credible and informative. 

